---
title: "LatNet analysis"
output: html_notebook
---

# load test data from the CoRegNet package
```{r}
library(CoRegNet)
data("CIT_BLCA_EXP")
normal_samples <- which(grepl('^N', colnames(CIT_BLCA_EXP)))
cancerExp <- CIT_BLCA_EXP[,-normal_samples]
dim(cancerExp)
```

# load LatNet functions (YOU SHOULD SET YOUR OWN PATH)
```{r}
LatNet_Path <- '~/Documents/GitHub/LatNet/LatNet.R'
source(LatNet_Path)
```

# Infer regulatory network for the genes. You can load your network structure directly if it has been obtained with different means 
```{r}
# perform the inference as discribed in the coregnet package
# Load list of Human transcription factors
data(HumanTF)
 
# you can set the option parallel="multicore" to accelerate the inference in case you have a multicore computing infrastructure
# Generate the network
GRN <- hLICORN(numericalExpression=cancerExp,
               TFlist = HumanTF,
               # minGeneSupport = 0.2,
               # minCoregSupport=0.2,
               maxCoreg = 2,
               verbose = TRUE,
               parallel = "multicore")
```

generating influences
```{r}
influences <- .regulatorActivity(GRN, cancerExp)
# inf <- regulatorInfluence(GRN, cancerExp)
```

heatmap with influences
```{r}
heatplot(influences[,names(CIT_BLCA_Subgroup)], groups = CIT_BLCA_Subgroup)
heatplot(inf[,names(CIT_BLCA_Subgroup)], groups = CIT_BLCA_Subgroup)
```

generating perturbations
```{r}
refData <- cancerExp
targetData <- cancerExp
ptm <- proc.time()
allperturbations.raw <- t(simplify2array(mclapply(targets(GRN), .oneGenePerturbations, GRN, t(refData), t(targetData), 0)))#, mc.cores = 3
rownames(allperturbations.raw) <- targets(GRN)
nrow(allperturbations.raw)
#remove the one having 0 everywhere
allperturbations <- allperturbations.raw[-which(rowSums(allperturbations.raw)==0),]
runtime <- proc.time() - ptm
print(runtime)
nrow(allperturbations)
```

Export all data if necessary
```{r}

ds1 <- t(cancerExp[c(targets(GRN), names(regulators(GRN))),names(CIT_BLCA_Subgroup)])
ds1 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds1)

ds2 <- t(influences[,names(CIT_BLCA_Subgroup)])
ds2 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds2)

ds3 <- t(allperturbations[,names(CIT_BLCA_Subgroup)])
ds3 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds3)

write.csv(ds1, file = "/Users/dhifli/UsexpressionsAllCIT.csv")
write.csv(ds2, file = "/Users/dhifli/influencesAllCIT.csv")
write.csv(ds3, file = "/Users/dhifli/perturbationsAllCIT.csv")
```

heatmap with perturbations
```{r}
heatplot(allperturbations[,names(CIT_BLCA_Subgroup)], groups = CIT_BLCA_Subgroup)
```

get only basal and luminals from Coregnet package
```{r}
data("CIT_BLCA_Subgroup")
annot_basal_luminal <- CIT_BLCA_Subgroup#[CIT_BLCA_Subgroup %in% c("basal-like", "luminal-1", "luminal-2")]

cit_exp_basal_luminal <- cancerExp#[,names(annot_basal_luminal)]
```

heatmap of the expression
```{r}
require(RColorBrewer)
mypalette<-RColorBrewer::brewer.pal(length(unique(annot_basal_luminal)),"Set3")
annot_colors <- mypalette[as.numeric(annot_basal_luminal)]

.distfun=function(x) as.dist((1 - cor(t(x))) )
# .distfun= function(x) dist(x,method="maximum")
.hclustfun = function(d) hclust(d,method="ward.D2")
# .hclustfun = function(d) hclust(d,method="complete")

par(mfrow=c(1,2))
heatmap.2(t(cit_exp_basal_luminal), distfun=.distfun,hclustfun=.hclustfun, RowSideColors = annot_colors
          , trace="none", cexRow=0.5
          , scale="column", dendrogram="col", Rowv = FALSE
          #, breaks= quantile(as.vector(t(cit_exp_basal_luminal)),probs=seq(0,1,by=1/75))
          )
par(lend = 1)
legend("topright",      # location of the legend on the heatmap plot
    legend = unique(annot_basal_luminal), # category labels
    col = unique(annot_colors),  # color key
    lty= 1,             # line style
    lwd = 10            # line width
)
```


heatmap of the influence
```{r}
require(RColorBrewer)
mypalette<-RColorBrewer::brewer.pal(length(unique(annot_basal_luminal)),"Set3")
annot_colors <- mypalette[as.numeric(annot_basal_luminal)]

.distfun=function(x) as.dist((1 - cor(t(x))) )
# .distfun= function(x) dist(x,method="euclidean")
.hclustfun = function(d) hclust(d,method="ward.D2")
# .hclustfun = function(d) hclust(d,method="centroid")

par(mfrow=c(1,2))
heatmap.2(t(influences[,names(annot_basal_luminal)]), distfun=.distfun,hclustfun=.hclustfun, RowSideColors = annot_colors
          , trace="none", cexRow=0.5
          # , scale="column"
          , dendrogram="col", Rowv = FALSE
          , breaks= unique(quantile(as.vector(t(influences[,names(annot_basal_luminal)])),probs=seq(0,1,by=1/75)))
          )
par(lend = 1)
legend("topright",      # location of the legend on the heatmap plot
    legend = unique(annot_basal_luminal), # category labels
    col = unique(annot_colors),  # color key
    lty= 1,             # line style
    lwd = 10            # line width
)
```


heatmap of the perturbations
```{r}
require(RColorBrewer)
mypalette<-RColorBrewer::brewer.pal(length(unique(annot_basal_luminal)),"Set3")
annot_colors <- mypalette[as.numeric(annot_basal_luminal)]

topN <- 30
g1 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])]), decreasing = TRUE)), topN)
g2 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])]), decreasing = TRUE)), topN)
g3 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])]), decreasing = TRUE)), topN)
g4 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])]), decreasing = TRUE)), topN)

top_p_genes <- unique(c(g1,g2,g3,g4))
top_p_genes <- names(which(rowMeans(abs(allperturbations))>=quantile(rowMeans(abs(allperturbations)))["75%"]))

.distfun=function(x) as.dist((1 - cor(t(x))) )
# .distfun= function(x) dist(x,method="maximum")
.hclustfun = function(d) hclust(d,method="ward.D2")
# .hclustfun = function(d) hclust(d,method="complete")

par(mfrow=c(1,2))
heatmap.2(t(allperturbations[top_p_genes,names(annot_basal_luminal)]), distfun=.distfun,hclustfun=.hclustfun, RowSideColors = annot_colors
          , trace="none", cexRow=0.5
          # , scale="column"
          # , dendrogram="col", Rowv = FALSE
          # , breaks= quantile(as.vector(t(allperturbations.raw[,names(annot_basal_luminal)])),probs=seq(0,1,by=1/75))
          )
par(lend = 1)
legend("topright",      # location of the legend on the heatmap plot
    legend = unique(annot_basal_luminal), # category labels
    col = unique(annot_colors),  # color key
    lty= 1,             # line style
    lwd = 10            # line width
)
```

classification using cca with the expression
```{r}
#CIT_BLCA_Subgroup[CIT_BLCA_Subgroup %in% c("basal-like", "luminal-1", "luminal-2")]
cancer_luminal1 <- cancerExp[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-1"])]
cancer_luminal2 <- cancerExp[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-2"])]
cancer_basal <- cancerExp[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="basal-like"])]

library(mixOmics)
#plot rcca
linn.res <- rcc(cancer_luminal1, cancer_luminal2)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal1, cancer_basal)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal2, cancer_basal)
plotVar(linn.res)
```


classification using cca with the influence
```{r}
#CIT_BLCA_Subgroup[CIT_BLCA_Subgroup %in% c("basal-like", "luminal-1", "luminal-2")]
cancer_luminal1 <- influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-1"])]
cancer_luminal2 <- influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-2"])]
cancer_basal <- influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="basal-like"])]

library(mixOmics)
#plot rcca
linn.res <- rcc(cancer_luminal1, cancer_luminal2)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal1, cancer_basal)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal2, cancer_basal)
plotVar(linn.res)
```

classification using cca with the perturbation
```{r}
#CIT_BLCA_Subgroup[CIT_BLCA_Subgroup %in% c("basal-like", "luminal-1", "luminal-2")]
cancer_luminal1 <- allperturbations[top_p_genes,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-1"])]
cancer_luminal2 <- allperturbations[top_p_genes,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="luminal-2"])]
cancer_basal <- allperturbations[top_p_genes,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup =="basal-like"])]

library(mixOmics)
#plot rcca
linn.res <- rcc(cancer_luminal1, cancer_luminal2)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal1, cancer_basal)
plotVar(linn.res)

linn.res <- rcc(cancer_luminal2, cancer_basal)
plotVar(linn.res)
```

classification using classifiers with the expression
```{r}
ds1 <- t(CIT_BLCA_EXP[c(targets(GRN), names(regulators(GRN))),names(CIT_BLCA_Subgroup)])
ds1 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds1)

ds2 <- t(influences[,names(CIT_BLCA_Subgroup)])
ds2 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds2)

ds3 <- t(allperturbations[,names(CIT_BLCA_Subgroup)])
ds3 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds3)

library(caret)
fitControl <- trainControl(## 5-fold CV
                           method = "cv",
                           number = 5
                           # ,summaryFunction = prSummary
                          )

set.seed(825)
model1 <- train(subtype ~ ., data = ds1, 
                 method = "svmRadial", 
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = TRUE)

model2 <- train(subtype ~ ., data = ds2, 
                 method = "svmRadial", 
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = TRUE)

model3 <- train(subtype ~ ., data = ds3, 
                 method = "svmRadial", 
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = TRUE)

```

classification using classifiers with the expression
```{r}
ds1 <- t(exp[c(targets(GRN), names(regulators(GRN))),names(CIT_BLCA_Subgroup)])#c(rownames(influences), rownames(allperturbations))
ds1 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds1)

library(ggplot2)
library(MASS)
library(caret)
lda1 <- lda(formula = subtype ~ ., data = ds1)
ldapred1 <- predict(lda1, ds1)
ggplot(as.data.frame(cbind(ldapred1$x[,"LD1"],ldapred1$x[,"LD2"],ldapred1$class)), aes (x = ldapred1$x[,"LD1"], y = ldapred1$x[,"LD2"], colour = ldapred1$class)) + geom_point(aes(ldapred1$x[,"LD1"], ldapred1$x[,"LD2"], colour = ldapred1$class, shape = ds1$subtype), size = 2.5) + labs(x=paste("LD1 (66%)"), y=paste("LD2 (26%)"), colour = "Predicted Subtype", shape = "True Subtype", title = "Gene Expression Signals") + stat_density2d ()
confusionMatrix(ldapred1$class, ds1$subtype)
# library(ModelMetrics)
# real <- ds1$subtype
# prediction <- ldapred1$class
# paste0("number of errors: ",ce(real, prediction)*length(real))
# paste0("Accuracy: ", 1-ce(real, prediction))
# paste0("F1-score: ", f1Score(real, prediction))
# paste0("Precision (postive predictive value): ", ppv(real, prediction))
# paste0("NPV (negative predictive value): ", npv(real, prediction))
# paste0("Sensitivity (recall, true positive rate): ", recall(real, prediction))
# paste0("Specificity (true negative rate): ", tnr(real, prediction))
```

classification using classifiers with the influence
```{r}

ds2 <- t(influences[,names(CIT_BLCA_Subgroup)])
ds2 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds2)

library(ggplot2)
library(MASS)
library(caret)

lda2 <- lda(formula = subtype ~ ., data = ds2)
ldapred2 <- predict(lda2, ds2)
ggplot(as.data.frame(cbind(ldapred2$x[,"LD1"],ldapred2$x[,"LD2"],ldapred2$class)), aes (x = ldapred2$x[,"LD1"], y = ldapred2$x[,"LD2"], colour = ldapred2$class)) + geom_point(aes(ldapred2$x[,"LD1"], ldapred2$x[,"LD2"], colour = ldapred2$class, shape = ds2$subtype), size = 2.5) + labs(x=paste("LD1 (63%)"), y=paste("LD2 (26%)"), colour = "Predicted Subtype", shape = "True Subtype", title = "Regulator Activity Signals") + stat_density2d ()
confusionMatrix(ldapred2$class, ds2$subtype)
# real <- ds2$subtype
# prediction <- ldapred2$class
# paste0("number of errors: ",ce(real, prediction)*length(real))
# paste0("Accuracy: ", 1-ce(real, prediction))
# paste0("F1-score: ", f1Score(real, prediction))
# paste0("Precision (postive predictive value): ", ppv(real, prediction))
# paste0("NPV (negative predictive value): ", npv(real, prediction))
# paste0("Sensitivity (recall, true positive rate): ", recall(real, prediction))
# paste0("Specificity (true negative rate): ", tnr(real, prediction))
```

classification using classifiers with the perturbation
```{r}
# topN <- 50
# g1 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])]), decreasing = TRUE)), topN)
# g2 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])]), decreasing = TRUE)), topN)
# g3 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])]), decreasing = TRUE)), topN)
# g4 <- head(names(sort(rowMeans(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])]), decreasing = TRUE)), topN)
# 
# top_p_genes <- unique(c(g1,g2,g3,g4))
# top_p_genes <- head(names(sort(rowSums(allperturbations), decreasing = TRUE)), 200)
top_p_genes <- names(which(rowMeans(abs(allperturbations))>=quantile(rowMeans(abs(allperturbations)))["75%"]))

ds3 <- t(allperturbations[top_p_genes,names(CIT_BLCA_Subgroup)])
ds3 <- data.frame("subtype"=CIT_BLCA_Subgroup, ds3)

library(ggplot2)
library(MASS)
library(caret)

lda3 <- lda(formula = subtype ~ ., data = ds3)
ldapred3 <- predict(lda3, ds3)
ggplot(as.data.frame(cbind(ldapred3$x[,"LD1"],ldapred3$x[,"LD2"],ldapred3$class)), aes (x = ldapred3$x[,"LD1"], y = ldapred3$x[,"LD2"], colour = ldapred3$class)) + geom_point(aes(ldapred3$x[,"LD1"], ldapred3$x[,"LD2"], colour = ldapred3$class, shape = ds3$subtype), size = 2.5) + labs(x=paste("LD1 (54%)"), y=paste("LD2 (27%)"), colour = "Predicted Subtype", shape = "True Subtype", title = "Gene Perturbation Signals") + stat_density2d ()
confusionMatrix(ldapred3$class, ds3$subtype)
# real <- ds3$subtype
# prediction <- ldapred3$class
# paste0("number of errors: ",ce(real, prediction)*length(real))
# paste0("Accuracy: ", 1-ce(real, prediction))
# paste0("F1-score: ", f1Score(real, prediction))
# paste0("Precision (postive predictive value): ", ppv(real, prediction))
# paste0("NPV (negative predictive value): ", npv(real, prediction))
# paste0("Sensitivity (recall, true positive rate): ", recall(real, prediction))
# paste0("Specificity (true negative rate): ", tnr(real, prediction))
```



classification using nearest shrunken centroids with the expression
```{r}
library(pamr)
khan.data <- list(x=exp[c(targets(GRN), names(regulators(GRN))),names(CIT_BLCA_Subgroup)],y=CIT_BLCA_Subgroup)

khan.data <- list(x=influences[,names(CIT_BLCA_Subgroup)],y=CIT_BLCA_Subgroup)

top_p_genes <- names(which(rowMeans(abs(allperturbations))>=quantile(rowMeans(abs(allperturbations)))["75%"]))
khan.data <- list(x=allperturbations[top_p_genes,names(CIT_BLCA_Subgroup)],y=CIT_BLCA_Subgroup)

# # interactive analysis
# pamr.menu(khan.data)

## Train the classifier
khan.trn <- pamr.train(khan.data)
## look for the threshold
khan.scales <- pamr.adaptthresh(khan.trn)
khan.train <- pamr.train(khan.data, threshold.scale=khan.scales)

## Type name of object to see the results
khan.train

## Cross-validate the classifier
khan.results<- pamr.cv(khan.train, khan.data)
khan.results

## Plot the cross-validated error curves
pamr.plotcv(khan.results)

thresh=0.3

## Compute the confusion matrix for a particular model (threshold=4.0) 
pamr.confusion(khan.results, threshold=thresh)

## Plot the cross-validated class probabilities by class
pamr.plotcvprob(khan.results, khan.data, threshold=thresh)

## Plot the class centroids
pamr.plotcen(khan.train, khan.data, threshold=thresh)

## Make a gene plot of the most significant genes
pamr.geneplot(khan.train, khan.data, threshold=thresh)


# Estimate false discovery rates and plot them
fdr.obj<- pamr.fdr(khan.train, khan.data)

pamr.plotfdr(fdr.obj)

## List the significant genes
pamr.listgenes(khan.train, khan.data, threshold=thresh)
```


Per subtype raking - influence
```{r}
topN <- 10
r1 <- head(names(sort(rowMeans(influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])]), decreasing = TRUE)), topN)
r2 <- head(names(sort(rowMeans(influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])]), decreasing = TRUE)), topN)
r3 <- head(names(sort(rowMeans(influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])]), decreasing = TRUE)), topN)
r4 <- head(names(sort(rowMeans(influences[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])]), decreasing = TRUE)), topN)
paste0("top-",topN," regulators for ",unique(CIT_BLCA_Subgroup)[1], " are : " )
paste0(r1)
paste0("top-",topN," regulators for ",unique(CIT_BLCA_Subgroup)[2], " are : " )
paste0(r2)
paste0("top-",topN," regulators for ",unique(CIT_BLCA_Subgroup)[3], " are : " )
paste0(r3)
paste0("top-",topN," regulators for ",unique(CIT_BLCA_Subgroup)[4], " are : " )
paste0(r4)

#overall perturbations
head(names(sort(rowMeans(influences), decreasing = TRUE)), topN)
```


Per subtype raking - perturbation
```{r}
topN <- 10
g1 <- head(names(sort(rowMeans(abs(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])])), decreasing = TRUE)), topN)
g2 <- head(names(sort(rowMeans(abs(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])])), decreasing = TRUE)), topN)
g3 <- head(names(sort(rowMeans(abs(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])])), decreasing = TRUE)), topN)
g4 <- head(names(sort(rowMeans(abs(allperturbations[,names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])])), decreasing = TRUE)), topN)
paste0("top-",topN," genes for ",unique(CIT_BLCA_Subgroup)[1], " are : " )
paste0(g1)
paste0("top-",topN," genes for ",unique(CIT_BLCA_Subgroup)[2], " are : " )
paste0(g2)
paste0("top-",topN," genes for ",unique(CIT_BLCA_Subgroup)[3], " are : " )
paste0(g3)
paste0("top-",topN," genes for ",unique(CIT_BLCA_Subgroup)[4], " are : " )
paste0(g4)

#overall perturbations
head(names(sort(rowMeans(abs(allperturbations)), decreasing = TRUE)), topN)
```

comparative network analysis Influence
```{r}
# #create the network
# library("psych")
library("qgraph")

luminal1_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])
luminal2_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])
basallike_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])
tcga4_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])



cormatInf <- cor(t(influences[,names(CIT_BLCA_Subgroup)]))
diag(cormatInf) <- 0
cormatInf[cormatInf<0.75] <- 0

# , minimum = 0.1, cut = 0.95
# qgraph(cormatInf, layout = "spring", scores = rowMeans(influences)[V(g1)], color="red", vsize=8, repulsion = 0.55)

qgInf <- qgraph(cormatInf, layout = "spring", repulsion = 0.55, vsize=8, color="red", cut = 0.95, labels=rownames(cormatInf), theme="TeamFortress")

par(mfrow=c(2,2))


qgraph(qgInf, scores = rowMeans(influences[,luminal1_refs])[qgInf$graphAttributes$Nodes$labels])
title("Luminal-1", line = 2, cex.main = 2.5)

qgraph(qgInf, scores = rowMeans(influences[,luminal2_refs])[qgInf$graphAttributes$Nodes$labels])
title("Luminal-2", line = 2, cex.main = 2.5)

qgraph(qgInf, layout = "spring", scores = rowMeans(influences[,basallike_refs])[qgInf$graphAttributes$Nodes$labels])
title("Basal-like", line = 2, cex.main = 2.5)

qgraph(qgInf, layout = "spring", scores = rowMeans(influences[,tcga4_refs])[qgInf$graphAttributes$Nodes$labels])
title("TCGA-4", line = 2, cex.main = 2.5)

# library("ppcor")
# pcormatInf <- pcor(t(influences[,names(CIT_BLCA_Subgroup)]))
# diag(pcormatInf) <- 0
# pcormatInf[pcormatInf<0.85] <- 0

# corGraph <- qgraph(cormatInf, layout = "spring", graph = "pcor", nodeNames = rownames(cormatInf), cut = 0.3, maximum = 1, minimum = 0, esize = 20, legend.cex = 0.3, vsize = 3, repulsion = 0.8)#, groups = Groups
# 
# library(igraph)
# g1 <- graph.adjacency(cormatInf, mode = "lower", weighted = TRUE)
# vcount(g1)
# ecount(g1)
# 
# V(g1)$size <- rowMeans(influences)[V(g1)]*3
# par(mfrow=c(1,2))
# # plot(g1,vertex.size=5)
# plot.igraph(g1, layout=layout.lgl(g1,repulserad=0.1), vertex.label.cex=1, vertex.label.degree=-pi/2)
# 
# plot.igraph(g1, layout=layout.fruchterman.reingold(g1, niter = 1000, area=10*vcount(g1)^2), vertex.label.cex=1, vertex.label.dist=0.5)
# 
# qgraph(cormatInf, layout = "spring", graph = "pcor", cut = 0.3, esize = 20, vsize = rowMeans(influences)[V(g1)], repulsion = 0.8)#, nodeNames = rownames(cormatInf)
# plot(g1,layout=l)

# qgraph(cormatInf, minimum = 0.9, cut = 0.9, layout = "spring", scores = rowMeans(influences)[V(g1)], repulsion = 0.4, vsize=6, color="red")
# # palf <- colorRampPalette(c(rgb(1,1,1, .2),rgb(.8,0,0, .7)), alpha=TRUE)
# 
# 
# layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 
# layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree|with_dh|randomly|kk|grid", layouts)]
# 
# par(mfrow=c(3,3), mar=c(1,1,1,1))
# 
# for (layout in layouts) {
# 
#   print(layout)
# 
#   l <- do.call(layout, list(g1)) 
# 
#   plot(g1, edge.arrow.mode=0, layout=l, main=layout) }


```


comparative network analysis Perturbation tests
```{r}
# #create the network
# library("psych")
library("qgraph")

luminal1_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])
luminal2_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])
basallike_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])
tcga4_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])

allperturbations_scaled <- scale(abs(allperturbations))
library(DMwR)
# allperturbations_scaled <- SoftMax(abs(allperturbations)))

# allperturbations_scaled[allperturbations_scaled>0] <- exp(allperturbations_scaled[allperturbations_scaled>0])
# allperturbations_scaled[allperturbations_scaled<0] <- -exp(abs(allperturbations_scaled[allperturbations_scaled<0]))
# allperturbations_scaled <- scale(allperturbations_scaled)


# allperturbations_scaled <- scale((abs(allperturbations)))

# 
topN <- 25
g1 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal1_refs])), decreasing = TRUE)), topN)
g2 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal2_refs])), decreasing = TRUE)), topN)
g3 <- head(names(sort(rowMeans((allperturbations_scaled[,basallike_refs])), decreasing = TRUE)), topN)
g4 <- head(names(sort(rowMeans((allperturbations_scaled[,tcga4_refs])), decreasing = TRUE)), topN)
top_p_genes <- unique(c(g1,g2,g3,g4))
# top_p_genes <- head(names(which(rowMeans((allperturbations_scaled))>=quantile(rowMeans((allperturbations_scaled)))["75%"])), 100)

cormatpert <- cor(t(allperturbations_scaled[top_p_genes,names(CIT_BLCA_Subgroup)]))
diag(cormatpert) <- 0
cormatpert[cormatpert<0.1] <- 0

# groups <- rep("A",length(rownames(cormatpert)))
# names(groups) <- rownames(cormatpert)
# groups[which(rowMeans(allperturbations_scaled[top_p_genes,])<0)] <- "B"
qgpert <- qgraph(cormatpert, layout = "spring", scores = (rowMeans(allperturbations_scaled[top_p_genes,])[rownames(cormatpert)]),
                 labels=rownames(cormatpert), theme="TeamFortress", vsize=8, color="darkorange", repulsion = 1)#, groups=as.vector(groups), cut = 0.95

par(mfrow=c(2,2))

qgraph(qgpert, scores = rowMeans((allperturbations_scaled[top_p_genes,luminal1_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-1", line = 2, cex.main = 2.5)

qgraph(qgpert, scores = rowMeans((allperturbations_scaled[top_p_genes,luminal2_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-2", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowMeans((allperturbations_scaled[top_p_genes,basallike_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Basal-like", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowMeans((allperturbations_scaled[top_p_genes,tcga4_refs]))[qgpert$graphAttributes$Nodes$labels])
title("TCGA-4", line = 2, cex.main = 2.5)


```

comparative network analysis Perturbation
```{r}
# #create the network
# library("psych")
library("qgraph")

luminal1_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])
luminal2_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])
basallike_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])
tcga4_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])

allperturbations_scaled <- allperturbations
# allperturbations_scaled[allperturbations_scaled>0] <- exp(allperturbations_scaled[allperturbations_scaled>0])
# allperturbations_scaled[allperturbations_scaled<0] <- -exp(abs(allperturbations_scaled[allperturbations_scaled<0]))
# allperturbations_scaled <- scale(allperturbations_scaled)


# allperturbations_scaled <- scale((abs(allperturbations)))

# 
topN <- 50
g1 <- head(names(sort(rowMeans(abs(allperturbations_scaled[,luminal1_refs])), decreasing = TRUE)), topN)
g2 <- head(names(sort(rowMeans(abs(allperturbations_scaled[,luminal2_refs])), decreasing = TRUE)), topN)
g3 <- head(names(sort(rowMeans(abs(allperturbations_scaled[,basallike_refs])), decreasing = TRUE)), topN)
g4 <- head(names(sort(rowMeans(abs(allperturbations_scaled[,tcga4_refs])), decreasing = TRUE)), topN)
top_p_genes <- unique(c(g1,g2,g3,g4))
# top_p_genes <- head(names(which(rowMeans((allperturbations_scaled))>=quantile(rowMeans((allperturbations_scaled)))["75%"])), 100)

cormatpert <- cor(t(allperturbations_scaled[top_p_genes,names(CIT_BLCA_Subgroup)]))
diag(cormatpert) <- 0
cormatpert[cormatpert<0.2] <- 0

# groups <- rep("A",length(rownames(cormatpert)))
# names(groups) <- rownames(cormatpert)
# groups[which(rowMeans(allperturbations_scaled[top_p_genes,])<0)] <- "B"
qgpert <- qgraph(cormatpert, layout = "spring", scores = (rowMeans(allperturbations_scaled[top_p_genes,])[rownames(cormatpert)]),
                 labels=rownames(cormatpert), theme="TeamFortress", vsize=8, color="darkorange", repulsion = 1)#, groups=as.vector(groups), cut = 0.95

par(mfrow=c(2,2))

qgraph(qgpert, scores = rowMeans(abs(allperturbations_scaled[top_p_genes,luminal1_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-1", line = 2, cex.main = 2.5)

qgraph(qgpert, scores = rowMeans(abs(allperturbations_scaled[top_p_genes,luminal2_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-2", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowMeans(abs(allperturbations_scaled[top_p_genes,basallike_refs]))[qgpert$graphAttributes$Nodes$labels])
title("Basal-like", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowMeans(abs(allperturbations_scaled[top_p_genes,tcga4_refs]))[qgpert$graphAttributes$Nodes$labels])
title("TCGA-4", line = 2, cex.main = 2.5)


```


comparative network analysis Perturbation by groups
```{r}
# #create the network
# library("psych")
library("qgraph")

luminal1_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])
luminal2_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])
basallike_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])
tcga4_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])

allperturbations_scaled <- scale(allperturbations)
# allperturbations_scaled[allperturbations_scaled>0] <- exp(allperturbations_scaled[allperturbations_scaled>0])
# allperturbations_scaled[allperturbations_scaled<0] <- -exp(abs(allperturbations_scaled[allperturbations_scaled<0]))
# allperturbations_scaled <- scale(allperturbations_scaled)


# allperturbations_scaled <- scale(exp(abs(allperturbations)))
library(DMwR)
# allperturbations_scaled <- SoftMax((allperturbations))
# 
topN <- 25
g1 <- c(head(names(sort(rowMeans((allperturbations_scaled[,luminal1_refs])), decreasing = TRUE)), topN), head(names(sort(rowMeans((allperturbations_scaled[,luminal1_refs])), decreasing = FALSE)), topN))
g2 <- c(head(names(sort(rowMeans((allperturbations_scaled[,luminal2_refs])), decreasing = TRUE)), topN),head(names(sort(rowMeans((allperturbations_scaled[,luminal2_refs])), decreasing = FALSE)), topN))
g3 <- c(head(names(sort(rowMeans((allperturbations_scaled[,basallike_refs])), decreasing = TRUE)), topN),head(names(sort(rowMeans((allperturbations_scaled[,basallike_refs])), decreasing = FALSE)), topN))
g4 <- c(head(names(sort(rowMeans((allperturbations_scaled[,tcga4_refs])), decreasing = TRUE)), topN),head(names(sort(rowMeans((allperturbations_scaled[,tcga4_refs])), decreasing = FALSE)), topN))
# g1 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal1_refs])), decreasing = TRUE)), topN)
# g2 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal2_refs])), decreasing = TRUE)), topN)
# g3 <- head(names(sort(rowMeans((allperturbations_scaled[,basallike_refs])), decreasing = TRUE)), topN)
# g4 <- head(names(sort(rowMeans((allperturbations_scaled[,tcga4_refs])), decreasing = TRUE)), topN)
top_p_genes <- unique(c(g1,g2,g3,g4))
# top_p_genes <- head(names(which(rowMeans((allperturbations_scaled))>=quantile(rowMeans((allperturbations_scaled)))["75%"])), 100)

cormatpert <- cor((t(allperturbations_scaled[top_p_genes,names(CIT_BLCA_Subgroup)])))
diag(cormatpert) <- 0
cormatpert[cormatpert<0.2] <- 0

thresh <- 0
groups <- rep("A",length(rownames(cormatpert)))
names(groups) <- rownames(cormatpert)
groups[which(rowMeans(allperturbations_scaled[top_p_genes,])<thresh)] <- "B"
qgpert <- qgraph(cormatpert, layout = "spring", scores = (rowMeans(allperturbations_scaled[top_p_genes,])[rownames(cormatpert)]),
                 labels=rownames(cormatpert), theme="TeamFortress", vsize=8, repulsion = 1, groups=as.vector(groups), legend=FALSE)#, cut = 0.95, color="darkorange"

par(mfrow=c(2,2))

groups <- rep("A",length(rownames(cormatpert)))
names(groups) <- rownames(cormatpert)
groups[which(rowMeans(allperturbations_scaled[top_p_genes,luminal1_refs])<thresh)] <- "B"

qgraph(qgpert, scores = abs(rowMeans(allperturbations_scaled[top_p_genes,luminal1_refs])[qgpert$graphAttributes$Nodes$labels]))
title("Luminal-1", line = 2, cex.main = 2.5)

groups <- rep("A",length(rownames(cormatpert)))
names(groups) <- rownames(cormatpert)
groups[which(rowMeans(allperturbations_scaled[top_p_genes,luminal2_refs])<thresh)] <- "B"

qgraph(qgpert, scores = abs(rowMeans(allperturbations_scaled[top_p_genes,luminal2_refs])[qgpert$graphAttributes$Nodes$labels]))
title("Luminal-2", line = 2, cex.main = 2.5)


groups <- rep("A",length(rownames(cormatpert)))
names(groups) <- rownames(cormatpert)
groups[which(rowMeans(allperturbations_scaled[top_p_genes,basallike_refs])<thresh)] <- "B"

qgraph(qgpert, layout = "spring", scores = abs(rowMeans(allperturbations_scaled[top_p_genes,basallike_refs])[qgpert$graphAttributes$Nodes$labels]))
title("Basal-like", line = 2, cex.main = 2.5)


groups <- rep("A",length(rownames(cormatpert)))
names(groups) <- rownames(cormatpert)
groups[which(rowMeans(allperturbations_scaled[top_p_genes,tcga4_refs])<thresh)] <- "B"

qgraph(qgpert, layout = "spring", scores = abs(rowMeans(allperturbations_scaled[top_p_genes,tcga4_refs])[qgpert$graphAttributes$Nodes$labels]))
title("TCGA-4", line = 2, cex.main = 2.5)


```

comparative network analysis Perturbation
```{r}
# #create the network
# library("psych")
library("qgraph")

luminal1_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[1]])
luminal2_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[2]])
basallike_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[3]])
tcga4_refs <- names(CIT_BLCA_Subgroup[CIT_BLCA_Subgroup==unique(CIT_BLCA_Subgroup)[4]])

allperturbations_scaled <- allperturbations
# allperturbations_scaled[allperturbations_scaled>0] <- exp(allperturbations_scaled[allperturbations_scaled>0])
# allperturbations_scaled[allperturbations_scaled<0] <- -exp(abs(allperturbations_scaled[allperturbations_scaled<0]))
# allperturbations_scaled <- scale(allperturbations_scaled)

# allperturbations_scaled <- scale(exp(abs(allperturbations)))

topN <- 25
g1 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal1_refs])), decreasing = TRUE)), topN)
g2 <- head(names(sort(rowMeans((allperturbations_scaled[,luminal2_refs])), decreasing = TRUE)), topN)
g3 <- head(names(sort(rowMeans((allperturbations_scaled[,basallike_refs])), decreasing = TRUE)), topN)
g4 <- head(names(sort(rowMeans((allperturbations_scaled[,tcga4_refs])), decreasing = TRUE)), topN)
top_p_genes <- unique(c(g1,g2,g3,g4))
# top_p_genes <- head(names(which(rowMeans((allperturbations_scaled))>=quantile(rowMeans((allperturbations_scaled)))["75%"])), 100)

cormatpert <- cor(t(allperturbations_scaled[top_p_genes,names(CIT_BLCA_Subgroup)]))
diag(cormatpert) <- 0
cormatpert[cormatpert<0.2] <- 0

###########################correlation with the alterations
cit_cna <- read.table("/Users/dhifli/ownCloud/Latent_Net_Representations/data/CIT_Bladder_CNA.txt",sep = " ", header = TRUE, check.names = FALSE)
cit_cna[is.na(cit_cna)] <- 0
cit_cna_binary <- cit_cna
cit_cna_binary[cit_cna_binary!=1] <- 0
cit_cna_binary[cit_cna_binary==1] <- 1
missing <- setdiff(top_p_genes, rownames(cit_cna_binary))
if(length(missing)>0){
  missingmat <- matrix(0L, nrow = length(missing), ncol = length(colnames(cit_cna_binary)))
  rownames(missingmat) <- missing
  colnames(missingmat) <- colnames(cit_cna_binary)
  cit_cna_binary <- rbind(cit_cna_binary, missingmat)
}


# groups <- rep("A",length(rownames(cormatpert)))
# names(groups) <- rownames(cormatpert)
# groups[which(rowMeans(allperturbations_scaled[top_p_genes,])<0)] <- "B"
qgpert <- qgraph(cormatpert, layout = "spring", scores = (rowMeans(allperturbations_scaled[top_p_genes,])[rownames(cormatpert)]),
                 labels=rownames(cormatpert), theme="TeamFortress", vsize=8, color="darkorange", repulsion = 1)

par(mfrow=c(2,2))

qgraph(qgpert, scores = rowSums((cit_cna_binary[top_p_genes,intersect(colnames(cit_cna_binary),luminal1_refs)]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-1", line = 2, cex.main = 2.5)

qgraph(qgpert, scores = rowSums((cit_cna_binary[top_p_genes,intersect(colnames(cit_cna_binary),luminal2_refs)]))[qgpert$graphAttributes$Nodes$labels])
title("Luminal-2", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowSums((cit_cna_binary[top_p_genes,intersect(colnames(cit_cna_binary),basallike_refs)]))[qgpert$graphAttributes$Nodes$labels])
title("Basal-like", line = 2, cex.main = 2.5)

qgraph(qgpert, layout = "spring", scores = rowSums((cit_cna_binary[top_p_genes,intersect(colnames(cit_cna_binary),tcga4_refs)]))[qgpert$graphAttributes$Nodes$labels])
title("TCGA-4", line = 2, cex.main = 2.5)


```
