---
title: "LatNet analysis"
output: html_notebook
---

# Load test data from the CoRegNet package
```{r}
library(CoRegNet)
data("CIT_BLCA_EXP")
dim(CIT_BLCA_EXP)
```

# Load LatNet functions (YOU SHOULD SET YOUR OWN PATH)
```{r}
LatNet_Path <- '~/Documents/GitHub/LatNet/LatNet.R'
source(LatNet_Path)
```

# Infer regulatory network for the genes. You can load your network structure directly if it has been obtained with different means 
```{r}
# perform the inference as discribed in the coregnet package
# Load list of Human transcription factors
data(HumanTF)
 
# Generate the network 
# you can set the option parallel="multicore" to accelerate the inference in case you have a multicore computing infrastructure
GRN <- hLICORN(numericalExpression=CIT_BLCA_EXP,
               TFlist = HumanTF,
               minGeneSupport = 0.2,
               minCoregSupport=0.2,
               maxCoreg = 3,
               verbose = TRUE,
               parallel = "no")
```

# Generating regulator activity signals
```{r}
activity <- .regulatorActivity(GRN, CIT_BLCA_EXP)
```


# Generating perturbations for a particular target gene (for example UPK2)
```{r}
refData <- CIT_BLCA_EXP
targetData <- CIT_BLCA_EXP
Gene <- targets(GRN)[1]
Gene_perturbations <- .oneGenePerturbations(gene = Gene, network = network, refEXP = t(refData), targetEXP = t(targetData))
print(Gene)
print(Gene_perturbations)
```

# Generating perturbations
```{r}
refData <- CIT_BLCA_EXP
targetData <- CIT_BLCA_EXP
allperturbations.raw <- t(simplify2array(mclapply(targets(GRN), .oneGenePerturbations, GRN, t(refData), t(targetData), 0)))#, mc.cores = 3
rownames(allperturbations.raw) <- targets(GRN)
#remove the ones having 0 everywhere
allperturbations <- allperturbations.raw[-which(rowSums(allperturbations.raw)==0),]
```

# Export all data if necessary
```{r}
write.csv(activity, file = "/Users/dhifli/activities.csv")
write.csv(allperturbations, file = "/Users/dhifli/perturbations.csv")
```

